---
title: "Using M-HRF-LSS with fmrireg"
output: rmarkdown::html_vignette
vignette:
  engine: knitr::rmarkdown
  encoding: UTF-8
  eval: FALSE
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

The `mhrf_lss()` function provides a user-friendly interface to the M-HRF-LSS (Manifold-Guided HRF Estimation and Trial-wise Deconvolution) pipeline that integrates seamlessly with the `fmrireg` package infrastructure.

## Quick Start

```{r setup}
library(manifoldhrf)
library(fmrireg)
```

### Basic Usage

```{r basic}
# Load your fMRI data using fmrireg
dataset <- fmri_dataset(
  scans = "path/to/bold.nii.gz",
  mask = "path/to/mask.nii.gz",
  TR = 2,
  run_length = c(200, 200)
)

# Fit M-HRF-LSS model
fit <- mhrf_lss(
  formula = ~ hrf(condition),
  dataset = dataset,
  estimation = "both"  # Estimate both condition and trial-wise betas
)

# View results
print(fit)
```

## HRF Library Options

The `mhrf_lss()` function supports multiple ways to specify the HRF library:

### 1. Using Canonical fmrireg HRFs

```{r canonical}
# Use default canonical HRFs from fmrireg
fit_canonical <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  hrf_library = "canonical"  # Default
)
```

### 2. Using Custom fmrireg HRF Objects

```{r custom_hrf}
# Create custom HRF library from fmrireg objects
hrf_library <- list(
  HRF_SPMG1,      # SPM canonical
  HRF_SPMG2,      # SPM + temporal derivative
  HRF_SPMG3,      # SPM + temporal & dispersion derivatives
  HRF_GAMMA,      # Gamma function
  HRF_GAUSSIAN    # Gaussian
)

# Add lagged versions
hrf_library <- c(
  hrf_library,
  lapply(c(1, 2, 3), function(lag) lag_hrf(HRF_SPMG1, lag))
)

fit_custom <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  hrf_library = hrf_library
)
```

### 3. Using a Pre-computed HRF Matrix

```{r matrix_hrf}
# Load or create your own HRF matrix (p timepoints x N HRFs)
hrf_matrix <- matrix(rnorm(30 * 50), 30, 50)  # Example

fit_matrix <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  hrf_library = hrf_matrix
)
```

## Manifold Parameter Presets

The package provides convenient parameter presets:

```{r presets}
# Conservative: Fewer dimensions, more regularization
fit_conservative <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  manifold_params = "conservative"
)

# Balanced: Default trade-off (recommended)
fit_balanced <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  manifold_params = "balanced"
)

# Aggressive: More dimensions, less regularization
fit_aggressive <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  manifold_params = "aggressive"
)
```

## Advanced Options

### Custom Parameters

```{r custom_params}
# Fine-tune all parameters
custom_params <- list(
  m_manifold_dim_target = 5,         # Target manifold dimensions
  m_manifold_dim_min_variance = 0.95, # Min variance to retain
  k_local_nn_for_sigma = 10,         # Neighbors for bandwidth
  lambda_gamma = 0.01,               # Gamma regularization
  lambda_spatial_smooth = 0.5,       # Spatial smoothing
  num_neighbors_Lsp = 6              # Spatial neighbors
)

fit_custom_params <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  manifold_params = custom_params
)
```

### Spatial Smoothing

```{r spatial}
# Enable spatial smoothing by providing voxel coordinates
fit_spatial <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  spatial_info = dataset$mask,  # Will extract coordinates
  manifold_params = list(
    lambda_spatial_smooth = 1.0  # Increase smoothing
  )
)
```

### Robust Estimation

```{r robust}
# Enable outlier detection and downweighting
fit_robust <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  robust = TRUE
)
```

### Memory-Efficient Processing

```{r chunked}
# Process large datasets in chunks
fit_chunked <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  strategy = "chunked",
  nchunks = 20  # Process in 20 chunks
)
```

## Working with Results

### Extract Coefficients

```{r coef}
# Get condition-level betas
betas_condition <- coef(fit, type = "condition")

# Get trial-wise betas
betas_trial <- coef(fit, type = "trial")

# Get estimated HRF shapes
hrfs <- coef(fit, type = "hrf")
```

### Visualize Results

```{r plot}
# Plot HRFs for selected voxels
plot(fit, type = "hrfs", voxels = c(100, 500, 1000))

# Plot manifold diagnostics
plot(fit, type = "manifold")

# QC summary of HRF metrics
plot_qc_summary(fit)
```

### Generate QC Report

```{r qc}
# Generate comprehensive quality control report
generate_qc_report(
  manifold = fit$manifold,
  hrfs = fit$hrfs,
  xi_coordinates = fit$xi_coordinates,
  betas_condition = fit$betas_condition,
  betas_trial = fit$betas_trial,
  output_file = "mhrf_lss_qc.html"
)
```

## Integration with fmrireg Workflow

The M-HRF-LSS interface is designed to work seamlessly with existing fmrireg workflows:

```{r integration}
# 1. Create event model with multiple conditions
event_data <- data.frame(
  onset = c(10, 30, 50, 70, 90, 110),
  condition = rep(c("faces", "houses", "objects"), 2),
  run = rep(1:2, each = 3)
)

dataset$event_table <- event_data

# 2. Specify complex model with baseline
fit_complex <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  baseline_model = baseline_model(
    basis = "poly",
    degree = 2  # Quadratic drift
  ),
  estimation = "both"
)

# 3. Compare with standard fmrireg GLM
fit_standard <- fmri_lm(
  ~ hrf(condition),
  dataset = dataset,
  baseline_model = baseline_model(degree = 2)
)

# M-HRF-LSS provides voxel-specific HRFs and trial-wise estimates
# that standard GLM cannot capture
```

## Tips and Best Practices

1. **Start with presets**: Use `"balanced"` parameters for initial analyses
2. **Check QC reports**: Always generate and review QC reports
3. **Use spatial smoothing**: Improves stability for neighboring voxels
4. **Enable robust mode**: For datasets with motion artifacts
5. **Monitor memory**: Use chunked processing for >50k voxels
6. **Validate HRF library**: Ensure sufficient HRF diversity

## Troubleshooting

### Memory Issues
```{r memory}
# Estimate memory requirements
mem_req <- estimate_memory_requirements(
  n_timepoints = 300,
  n_voxels = 100000,
  n_conditions = 5,
  n_trials = 200
)
print(mem_req)

# Use more chunks if needed
fit <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  strategy = "chunked",
  nchunks = ceiling(mem_req$recommended_chunks)
)
```

### Convergence Issues
```{r convergence}
# Use conservative parameters
fit <- mhrf_lss(
  ~ hrf(condition),
  dataset = dataset,
  manifold_params = "conservative",
  robust = TRUE
)
```