---
title: "M-HRF-LSS Pipeline QC Report"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: cosmo
    highlight: tango
    code_folding: hide
    self_contained: true
params:
  results: NULL
  parameters: NULL
  metadata: NULL
  log: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6,
  fig.align = "center"
)

# Load required libraries
library(manifoldhrf)
library(ggplot2)
library(plotly)
library(DT)
library(knitr)

# Set theme
theme_set(theme_minimal())

# Extract parameters
results <- params$results
parameters <- params$parameters
metadata <- params$metadata
```

# Executive Summary

```{r summary}
# Calculate key QC metrics
n_voxels <- ncol(results$core_matrices$Y_data)
n_timepoints <- nrow(results$core_matrices$Y_data)
n_conditions <- nrow(results$core_matrices$Beta_condition_final)
n_trials <- nrow(results$core_matrices$Beta_trial)
manifold_dim <- ncol(results$core_matrices$Xi_smoothed)
n_truncated_hrfs <- results$qc_metrics$n_truncated_hrfs %||% 0

# QC flags
qc_flags <- list()

# Check for low trial count
if (n_trials < 20) {
  qc_flags$low_trial_count <- list(
    status = "warning",
    message = sprintf("Low trial count: %d trials (recommended: ≥20)", n_trials)
  )
}

# Calculate mean R² if available
if (!is.null(results$diagnostics$r2_voxelwise)) {
  mean_r2 <- mean(results$diagnostics$r2_voxelwise, na.rm = TRUE)
  poor_fit_fraction <- mean(results$diagnostics$r2_voxelwise < 0.1, na.rm = TRUE)
  
  if (poor_fit_fraction > 0.3) {
    qc_flags$poor_fits <- list(
      status = "warning",
      message = sprintf("%.1f%% of voxels have R² < 0.1", 100 * poor_fit_fraction)
    )
  }
} else {
  mean_r2 <- NA
  poor_fit_fraction <- NA
}

# Overall QC status
qc_status <- if (length(qc_flags) == 0) {
  "PASS"
} else if (any(sapply(qc_flags, function(x) x$status == "error"))) {
  "FAIL"
} else {
  "WARNING"
}

# Status badge color
status_color <- switch(qc_status,
  "PASS" = "success",
  "WARNING" = "warning",
  "FAIL" = "danger"
)
```

<div class="alert alert-`r status_color`" role="alert">
  <h4 class="alert-heading">QC Status: `r qc_status`</h4>
  <hr>
  <p><strong>Data dimensions:</strong> `r n_timepoints` timepoints × `r n_voxels` voxels</p>
  <p><strong>Experimental design:</strong> `r n_conditions` conditions, `r n_trials` trials</p>
  <p><strong>Manifold dimension:</strong> `r manifold_dim` (target: `r parameters$manifold$m_manifold_dim_target`)</p>
  <p><strong>Truncated HRFs:</strong> `r n_truncated_hrfs`</p>
  `r if (!is.na(mean_r2)) paste0("<p><strong>Mean voxel R²:</strong> ", sprintf("%.3f", mean_r2), "</p>")`
</div>

```{r qc_flags}
if (length(qc_flags) > 0) {
  cat("\n### QC Flags\n\n")
  for (flag_name in names(qc_flags)) {
    flag <- qc_flags[[flag_name]]
    badge_color <- ifelse(flag$status == "error", "danger", flag$status)
    cat(sprintf(
      '<span class="badge bg-%s">%s</span> %s\n\n',
      badge_color,
      gsub("_", " ", toupper(flag_name)),
      flag$message
    ))
  }
}
```

# Input Parameters

```{r parameters_table}
# Create parameters data frame
param_df <- data.frame(
  Parameter = character(),
  Value = character(),
  stringsAsFactors = FALSE
)

# Add manifold parameters
if (!is.null(parameters$manifold)) {
  for (name in names(parameters$manifold)) {
    param_df <- rbind(param_df, data.frame(
      Parameter = paste0("manifold.", name),
      Value = as.character(parameters$manifold[[name]])
    ))
  }
}

# Add pipeline parameters
if (!is.null(parameters$pipeline)) {
  for (name in names(parameters$pipeline)) {
    param_df <- rbind(param_df, data.frame(
      Parameter = paste0("pipeline.", name),
      Value = as.character(parameters$pipeline[[name]])
    ))
  }
}

# Display table
if (nrow(param_df) > 0) {
  datatable(param_df, 
            options = list(pageLength = 20, dom = 'ft'),
            rownames = FALSE)
}
```

# Manifold Diagnostics

## Eigenvalue Spectrum

```{r eigenvalue_spectrum, fig.height=5}
if (!is.null(results$manifold$eigenvalues_S)) {
  eigenvals <- results$manifold$eigenvalues_S[-1]  # Remove first trivial eigenvalue
  n_show <- min(50, length(eigenvals))
  
  df_eigen <- data.frame(
    Component = 1:n_show,
    Eigenvalue = eigenvals[1:n_show],
    Cumulative_Variance = cumsum(eigenvals[1:n_show]) / sum(eigenvals)
  )
  
  # Scree plot
  p1 <- ggplot(df_eigen, aes(x = Component, y = Eigenvalue)) +
    geom_line(size = 1.2, color = "steelblue") +
    geom_point(size = 3, color = "steelblue") +
    geom_vline(xintercept = manifold_dim, linetype = "dashed", color = "red") +
    scale_y_log10() +
    labs(title = "Eigenvalue Spectrum (Scree Plot)",
         subtitle = sprintf("Red line: Selected dimension (m = %d)", manifold_dim),
         x = "Component", y = "Eigenvalue (log scale)") +
    theme_minimal()
  
  print(p1)
}
```

## Variance Explained

```{r variance_explained, fig.height=5}
if (!is.null(results$manifold$eigenvalues_S)) {
  # Cumulative variance plot
  p2 <- ggplot(df_eigen, aes(x = Component, y = Cumulative_Variance)) +
    geom_line(size = 1.2, color = "darkgreen") +
    geom_point(size = 3, color = "darkgreen") +
    geom_hline(yintercept = parameters$manifold$m_manifold_dim_min_variance, 
               linetype = "dashed", color = "orange") +
    geom_vline(xintercept = manifold_dim, linetype = "dashed", color = "red") +
    scale_y_continuous(labels = scales::percent) +
    labs(title = "Cumulative Variance Explained",
         subtitle = sprintf("Orange: Target variance (%.0f%%), Red: Selected dimension",
                           100 * parameters$manifold$m_manifold_dim_min_variance),
         x = "Component", y = "Cumulative Variance") +
    theme_minimal()
  
  print(p2)
  
  # Report auto-selected dimension
  if (!is.null(results$manifold$m_auto_selected)) {
    cat(sprintf("\n**Auto-selected dimension:** %d (for %.0f%% variance)\n",
                results$manifold$m_auto_selected,
                100 * parameters$manifold$m_manifold_dim_min_variance))
    cat(sprintf("**Final dimension used:** %d\n", manifold_dim))
  }
}
```

## Reconstruction Error

```{r reconstruction_error, fig.height=5}
if (!is.null(results$diagnostics$reconstruction_error)) {
  df_recon <- data.frame(
    Dimension = 1:length(results$diagnostics$reconstruction_error),
    Error = results$diagnostics$reconstruction_error
  )
  
  p3 <- ggplot(df_recon, aes(x = Dimension, y = Error)) +
    geom_line(size = 1.2, color = "darkred") +
    geom_point(size = 3, color = "darkred") +
    geom_vline(xintercept = manifold_dim, linetype = "dashed", color = "red") +
    labs(title = "HRF Library Reconstruction Error",
         subtitle = "||L - BΦ||_F / ||L||_F vs manifold dimension",
         x = "Manifold Dimension", y = "Relative Reconstruction Error") +
    theme_minimal()
  
  print(p3)
}
```

# HRF Diagnostics

## Manifold Coordinate Maps

```{r xi_maps, fig.height=8}
# Show first 3 manifold coordinates
n_coords_show <- min(3, nrow(results$core_matrices$Xi_smoothed))

for (i in 1:n_coords_show) {
  xi_values <- results$core_matrices$Xi_smoothed[i, ]
  
  # Create simple visualization
  p <- ggplot(data.frame(voxel = 1:length(xi_values), value = xi_values),
               aes(x = voxel, y = value)) +
    geom_point(alpha = 0.5, size = 0.5) +
    geom_smooth(method = "loess", se = FALSE, color = "blue") +
    labs(title = sprintf("Manifold Coordinate %d (ξ_%d)", i, i),
         x = "Voxel Index", y = "Coordinate Value") +
    theme_minimal()
  
  print(p)
}
```

## Spatial Smoothing Effect

```{r smoothing_effect, fig.height=6}
if (!is.null(results$core_matrices$Xi_ident) && !is.null(results$core_matrices$Xi_smoothed)) {
  # Show effect of spatial smoothing on first coordinate
  coord_idx <- 1
  xi_before <- results$core_matrices$Xi_ident[coord_idx, ]
  xi_after <- results$core_matrices$Xi_smoothed[coord_idx, ]
  
  df_smooth <- data.frame(
    voxel = rep(1:length(xi_before), 2),
    value = c(xi_before, xi_after),
    type = rep(c("Before Smoothing", "After Smoothing"), each = length(xi_before))
  )
  
  p <- ggplot(df_smooth, aes(x = voxel, y = value, color = type)) +
    geom_point(alpha = 0.3, size = 0.5) +
    facet_wrap(~type, ncol = 1) +
    scale_color_manual(values = c("Before Smoothing" = "red", "After Smoothing" = "blue")) +
    labs(title = sprintf("Spatial Smoothing Effect on ξ_%d", coord_idx),
         x = "Voxel Index", y = "Coordinate Value") +
    theme_minimal() +
    theme(legend.position = "none")
  
  print(p)
}
```

## HRF Shape Statistics

```{r hrf_stats, fig.height=8}
if (!is.null(results$core_matrices$H_shapes)) {
  # Calculate HRF statistics
  n_voxels_hrf <- ncol(results$core_matrices$H_shapes)
  hrf_stats <- data.frame(
    voxel = 1:n_voxels_hrf,
    peak_time = NA,
    peak_amplitude = NA,
    fwhm = NA
  )
  
  # Time grid (assuming uniform sampling)
  time_points <- seq(0, (nrow(results$core_matrices$H_shapes) - 1) * parameters$manifold$TR_precision,
                     by = parameters$manifold$TR_precision)
  
  for (v in 1:n_voxels_hrf) {
    hrf <- results$core_matrices$H_shapes[, v]
    
    # Find peak
    peak_idx <- which.max(hrf)
    hrf_stats$peak_time[v] <- time_points[peak_idx]
    hrf_stats$peak_amplitude[v] <- hrf[peak_idx]
    
    # Estimate FWHM
    half_max <- max(hrf) / 2
    above_half <- which(hrf >= half_max)
    if (length(above_half) >= 2) {
      hrf_stats$fwhm[v] <- time_points[max(above_half)] - time_points[min(above_half)]
    }
  }
  
  # Peak time histogram
  p1 <- ggplot(hrf_stats, aes(x = peak_time)) +
    geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7) +
    geom_vline(xintercept = 5, linetype = "dashed", color = "red") +
    labs(title = "Distribution of HRF Peak Times",
         subtitle = "Red line: Canonical peak time (5s)",
         x = "Peak Time (seconds)", y = "Count") +
    theme_minimal()
  
  # FWHM histogram
  p2 <- ggplot(hrf_stats[!is.na(hrf_stats$fwhm), ], aes(x = fwhm)) +
    geom_histogram(bins = 30, fill = "darkgreen", alpha = 0.7) +
    labs(title = "Distribution of HRF Full Width at Half Maximum",
         x = "FWHM (seconds)", y = "Count") +
    theme_minimal()
  
  print(p1)
  print(p2)
  
  # Flag unusual HRFs
  unusual_peaks <- sum(hrf_stats$peak_time < 2 | hrf_stats$peak_time > 10, na.rm = TRUE)
  if (unusual_peaks > 0) {
    cat(sprintf("\n**Warning:** %d voxels (%.1f%%) have unusual peak times (<2s or >10s)\n",
                unusual_peaks, 100 * unusual_peaks / n_voxels_hrf))
  }
}
```

# Model Fit Diagnostics

## Voxel-wise R²

```{r r2_distribution, fig.height=6}
if (!is.null(results$diagnostics$r2_voxelwise)) {
  df_r2 <- data.frame(r2 = results$diagnostics$r2_voxelwise)
  
  p1 <- ggplot(df_r2, aes(x = r2)) +
    geom_histogram(bins = 50, fill = "darkblue", alpha = 0.7) +
    geom_vline(xintercept = median(df_r2$r2, na.rm = TRUE), 
               linetype = "dashed", color = "red", size = 1) +
    labs(title = "Distribution of Voxel-wise R²",
         subtitle = sprintf("Median R² = %.3f", median(df_r2$r2, na.rm = TRUE)),
         x = "R²", y = "Count") +
    theme_minimal()
  
  print(p1)
  
  # Spatial map of R² (simplified as line plot)
  p2 <- ggplot(data.frame(voxel = 1:length(results$diagnostics$r2_voxelwise),
                          r2 = results$diagnostics$r2_voxelwise),
               aes(x = voxel, y = r2)) +
    geom_point(alpha = 0.3, size = 0.5) +
    geom_smooth(method = "loess", se = FALSE, color = "blue") +
    geom_hline(yintercept = 0.1, linetype = "dashed", color = "red") +
    labs(title = "Voxel-wise R² Across Brain",
         subtitle = "Red line: R² = 0.1 threshold",
         x = "Voxel Index", y = "R²") +
    theme_minimal()
  
  print(p2)
}
```

## Convergence (Component 4)

```{r convergence, fig.height=5}
if (!is.null(results$diagnostics$convergence)) {
  df_conv <- data.frame(
    iteration = 1:length(results$diagnostics$convergence$beta_change),
    beta_change = results$diagnostics$convergence$beta_change,
    hrf_change = results$diagnostics$convergence$hrf_change
  )
  
  if (nrow(df_conv) > 1) {
    df_conv_long <- tidyr::pivot_longer(df_conv, cols = c(beta_change, hrf_change),
                                        names_to = "metric", values_to = "change")
    
    p <- ggplot(df_conv_long, aes(x = iteration, y = change, color = metric)) +
      geom_line(size = 1.2) +
      geom_point(size = 3) +
      scale_y_log10() +
      scale_color_manual(values = c("beta_change" = "blue", "hrf_change" = "red"),
                         labels = c("Beta Change", "HRF Change")) +
      labs(title = "Alternating Optimization Convergence",
           x = "Iteration", y = "Relative Change (log scale)") +
      theme_minimal() +
      theme(legend.title = element_blank())
    
    print(p)
  } else {
    cat("\nOnly 1 iteration performed (no convergence plot available)\n")
  }
}
```

# Performance Metrics

```{r performance}
if (!is.null(metadata$timing)) {
  timing_df <- data.frame(
    Component = names(metadata$timing),
    Time_seconds = unlist(metadata$timing)
  )
  
  timing_df$Time_minutes <- timing_df$Time_seconds / 60
  
  p <- ggplot(timing_df, aes(x = reorder(Component, Time_seconds), y = Time_minutes)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    coord_flip() +
    labs(title = "Computation Time by Component",
         x = "", y = "Time (minutes)") +
    theme_minimal()
  
  print(p)
  
  cat(sprintf("\n**Total processing time:** %.1f minutes\n", sum(timing_df$Time_minutes)))
}
```

# Pipeline Log

```{r pipeline_log}
if (!is.null(params$log)) {
  cat("<pre>\n")
  cat(paste(tail(params$log, 50), collapse = "\n"))
  cat("\n</pre>")
} else if (!is.null(results$log)) {
  cat("<pre>\n")
  cat(paste(tail(results$log, 50), collapse = "\n"))
  cat("\n</pre>")
} else {
  cat("No log available.")
}
```

# Failure Mode Checklist

Based on the QC flags and diagnostics, here are potential issues to investigate:

```{r failure_checklist}
checklist <- character()

# Low R²
if (!is.na(poor_fit_fraction) && poor_fit_fraction > 0.3) {
  checklist <- c(checklist, 
    "High proportion of poor fits - Check motion parameters and data quality",
    "Consider increasing regularization parameters or adjusting manifold dimension")
}

# Unusual HRFs
if (exists("unusual_peaks") && unusual_peaks > 0) {
  checklist <- c(checklist,
    "Unusual HRF peak times detected - Review HRF library and manifold construction",
    "Check TR specification and event timing")
}

# Low trial count
if (n_trials < 20) {
  checklist <- c(checklist,
    "Low trial count may affect LSS stability - Consider condition-level analysis only")
}

# Manifold dimension
if (!is.null(results$manifold$m_auto_selected) && 
    abs(manifold_dim - results$manifold$m_auto_selected) > 2) {
  checklist <- c(checklist,
    sprintf("Large discrepancy between target (m=%d) and auto-selected (m=%d) dimensions",
            manifold_dim, results$manifold$m_auto_selected),
    "Review eigenvalue spectrum and consider adjusting target dimension")
}

if (length(checklist) > 0) {
  cat("\n<div class='alert alert-info'>\n")
  for (item in checklist) {
    cat(sprintf("- %s\n", item))
  }
  cat("</div>\n")
} else {
  cat("\n<div class='alert alert-success'>")
  cat("No major issues detected. Pipeline completed successfully.")
  cat("</div>\n")
}
```

# Session Information

```{r session_info}
cat("\n```\n")
print(sessionInfo())
cat("```\n")
```